Quotient maps and configuration spaces of hard disks

### 2022-02-09 manuscript/main.tex
introduction paranthesis added to sigma_nu

Local minima are specific examples of a larger class of points known as critical points, roughly defined as locations where the topology of a manifold changes. The number and distribution of critical points of the potential energy surface has also been implicated in the onset of phase transitions, an idea known as the Topological Hypothesis \cite{caiani1997geometry,franzosi2004theorem}. Consider a system of particles with positions $\vec{q}_i$ and potential energy $V(\vec{q}_1,\dots,\vec{q}_N)$. Previously, Franzosi \etal \cite{franzosi2007topology1,franzosi2007topology2,gori2018topological} claimed that a change in the topology of the equipotential energy submanifolds $\Sigma_\nu = V^{-1}(-\infty, \nu]$ of the configuration space as a function of the energy $\nu$ is a necessary condition for a phase transition to occur in systems with smooth, stable, confining, and short-range interactions. Kastner and Mehta \cite{kastner2011phase} eventually found a counterexample satisfying all the stated conditions, but for which a phase transition occurs without a change in topology. They then proposed new criteria stating that a phase transition requires either (i) the number of critical points in a narrow potential energy band to grow exponentially faster than the number of particles, or (ii) the determinant of the Hessian matrix to vanish for a significant fraction of the critical points. It is significant that the Topological Hypothesis, either the original or the revised one, has so far only been evaluated for systems simple enough to be treated at least partially analytically; there appears to not yet even be the machinery available to test the hypothesis for, e.g., a simple fluid.

>>>>

Local minima are specific examples of a larger class of points known as critical points, roughly defined as locations where the topology of a manifold changes. The number and distribution of critical points of the potential energy surface has also been implicated in the onset of phase transitions, an idea known as the Topological Hypothesis \cite{caiani1997geometry,franzosi2004theorem}. Consider a system of particles with positions $\vec{q}_i$ and potential energy $V(\vec{q}_1,\dots,\vec{q}_N)$. Previously, Franzosi \etal \cite{franzosi2007topology1,franzosi2007topology2,gori2018topological} claimed that a change in the topology of the equipotential energy submanifolds $\Sigma_\nu = V^{-1}\left((-\infty, \nu]\right)$ of the configuration space as a function of the energy $\nu$ is a necessary condition for a phase transition to occur in systems with smooth, stable, confining, and short-range interactions. Kastner and Mehta \cite{kastner2011phase} eventually found a counterexample satisfying all the stated conditions, but for which a phase transition occurs without a change in topology. They then proposed new criteria stating that a phase transition requires either (i) the number of critical points in a narrow potential energy band to grow exponentially faster than the number of particles, or (ii) the determinant of the Hessian matrix to vanish for a significant fraction of the critical points. It is significant that the Topological Hypothesis, either the original or the revised one, has so far only been evaluated for systems simple enough to be treated at least partially analytically; there appears to not yet even be the machinery available to test the hypothesis for, e.g., a simple fluid.
